\LARGE \bf Geometric Adaptive Control of Attitude Dynamics on $\SO$\\ with State Inequality Constraints

### 2015-09-22 ACC16.2.tex
completed attitude control without disturbance

\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}

\end{prop}

@@

\subsection{Attitude Control without disturbance}
We introduce a nonlinear geometric controller for the attitude stabilization of a rigid body.
We assume that there is no disturbance, i.e., \( \Delta = 0 \).
\begin{prop}[Attitude Control]\label{prop:att_control}

\end{prop}
Introduce control input function and show that it is stable without any disturbance

Lyapunov function

@@

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined as \( Q \in SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \), namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180},
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \).

\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} A = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} B = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} e_{R_A} = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.

>>>>

\begin{prop}[Error Dynamics]\label{prop:error_dyn}
	The attitude error dynamics for \( \Psi, e_R, e_\Omega \) satisfy 
	\begin{gather}
		\diff{}{t} \parenth{\Psi} = e_R \cdot e_\Omega \, , \label{eqn:psi_dot}\\
		\diff{}{t} \parenth{e_R} = \dot{e}_{R_A} B + e_{R_A} \dot{B} + \dot{A}e_{R_B} + A \dot{e}_{R_B} \, , \label{eqn:eR_dot}\\
		\diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega \, , \label{eqn:eRA_dot} \\
		\diff{}{t} \parenth{e_{R_B}} = F(R) e_\Omega \, , \label{eqn:eRB_dot} \\
		\diff{}{t} \parenth{A} = e_{R_A} \cdot e_\Omega \, , \label{eqn:A_dot} \\
		\diff{}{t} \parenth{B} = e_{R_B} \cdot e_\Omega \, , \label{eqn:B_dot} \\
		\diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\Omega \times J \Omega + u + W(R, \Omega) \Delta} \, , \label{eqn:eW_dot}
	\end{gather}
	where the matrices \(E(R,R_d), F(R) \in \R^{3\times3} \) are given by
	\begin{gather}
		E(R,R_d) = \frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} \, , \label{eqn:E} \\
		F(R) = \frac{1}{\alpha \parenth{r^T R^T v - \cos \theta}} \left[\parenth{v^T R r} I - R^T v r^T + \right. \nonumber \\
		\left. \frac{R^T \hat{v} R r v^T R \hat{r}}{\parenth{r^T R^T v - \cos \theta}}\right] \, . \label{eqn:F}
	\end{gather}
\end{prop}

@@

\subsection{Attitude Control without disturbance}
We introduce a nonlinear geometric controller for the attitude stabilization of a rigid body.
We assume that there is no disturbance, i.e., \( \Delta = 0 \).
\begin{prop}[Attitude Control]\label{prop:att_control}
	Given a desired attitude command \( \parenth{R_d, \Omega_d = 0} \) and positive constants \( k_R, k_\Omega \in \R \) we define a control input \( u \in \R^3 \) as follows
	\begin{gather}
		u = -k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega \, . \label{eqn:nodist_control}
	\end{gather}
	Then the zero equilibrium of the attitude errors \( \parenth{e_R, e_\Omega} \) is asymptotically stable, i.e., \(e_R, e_\Omega \to 0 \) as \( t \to \infty \).
\end{prop}

\begin{proof}\label{proof:att_control}
	Consider the following Lyapunov function:
	\begin{gather}
		V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi(R,R_d) \, . \label{eqn:v_nodist}
	\end{gather}
	From~\cref{item:prop_psi_psd}, we obtain
	\begin{gather*}
		V \geq 0 \, .
	\end{gather*}
	Using~\cref{eqn:eW_dot,eqn:psi_dot} with \( \Delta = 0 \), the time derivative of \( V \) is given by
	\begin{align}
		\dot{V} &= e_\Omega^T \parenth{ u - \Omega \times J \Omega } + k_R e_R \cdot e_\Omega \nonumber \\
		&= -k_\Omega \norm{e_\Omega}^2 \, . \label{eqn:vdot_nodist}
	\end{align}
	Since \( V \) is positive definite and \( \dot{V} \) is negative definite this implies that the zero equilibrium point \( e_R, e_\Omega \) is asymptotically stable.
\end{proof}

This proposition only guarantees that the attitude error vector \( e_R \) asymptotically converges to zero.
However, this does not necessarily show that \( R \to R_d \) as \( t \to \infty \). 
There are three additional critical points of \( \Psi \) where \( e_R = 0 \).
At an undesired equilibrium \( R = \exp{\pi \hat{e}_i} R_d \) and \( e_\Omega =0 \). 
This results in \( A \) having a value of \( g_1 + g_2, g_2 + g_3 \) or \( g_3 + g_1 \).
At any of these undesired equilibria we have \( V = k \). 
We define an augmented Lyapunov function as
\begin{align*}
	W = k - V = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R (\frac{k}{k_R}+ \Psi(R,R_d)) \, .
\end{align*}
We can choose an attitude \( R \) arbitrarily close to the equilibrium point \( \exp{ \pi \hat{e}_i} R_d \) such that the \( k_r \parenth{\frac{k}{k_R}+\Psi }> 0 \).
This is due to the fact that \( \Psi \) is positive definite about \( R = R_d \) as shown in~\cref{item:prop_psi_psd}.
If \( e_\Omega \) is sufficiently small this results in \( W > 0 \) and \( \dot{W} = -\dot{V} > 0\) from~\cref{eqn:vdot_nodist}.
According to Chateav's theorem the undesired equilibria are unstable~\cite{khalil1996}.
As a result, we can claim that the desired equilibrium \( R = R_d \text{ and } e_\Omega = 0 \) is almost globally asymptotically stable.

@@

To show~\cref{item:prop_crit}, we determine the critical points of~\cref{eqn:eRA}.
The attitude error between \( R \) and \( R_d \) is defined as \( Q \in SO \) as
\begin{align*}
	Q = R_d^T R = 
	\begin{bmatrix}
		q_{11} & q_{12} & q_{13} \\
		q_{21} & q_{22} & q_{23} \\
		q_{31} & q_{32} & q_{33}
	\end{bmatrix} \, .
\end{align*}
Equating \( e_{R_{A}} = 0 \) yields the following equivalent relations
\begin{subequations}
\begin{align}
	g_3 q_{32} &= g_2 q_{23} \, , \\
	g_1 q_{13} &= g_3 q_{31} \, , \\
	g_2 q_{21} &= g_1 q_{12} \, .
\end{align}
\end{subequations}
These relations allow one to redefine the attitude error \( Q \) in terms of the independent upper diagonal elements.
Noting that \( Q \in \SO \) is an orthonormal matrix, with unit length row and columns gives
\begin{subequations}
\begin{align}
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 + \parenth{1 - \frac{g_1^2}{g_3^2}}q_{13}^2 &= 0 \, , \\
	\parenth{1 - \frac{g_1^2}{g_2^2}} q_{12}^2 - \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, , \\
	\parenth{1-\frac{g_1^2}{g_3^2}} q_{13}^2 + \parenth{1 - \frac{g_2^2}{g_3^2}} q_{23}^2 &= 0 \, .
\end{align}
\end{subequations}
As a result, all the off diagonal terms of \( Q = R_d^T R \), namely \( q_{12}= q_{13} = q_{23} = 0 \).
Since \( Q \in \SO \), the diagonal elements \( q_{11}, q_{22}, \text{ and } q_{33} \) are either \( +1 \) or \(-1\).
There are four critical points of \( e_{R_A} \), the desired attitude \( R_d \) as well as rotations about each body fixed axis by \ang{180}.
The repulsive error vector \( e_{R_{B}} \) is zero only when the numerator \( \parenth{R^T v}^\wedge r = 0 \). 
This condition only occurs if the desired attitude results in the body fixed vector \( r \) becoming aligned with \(R^T v \) while simultaneously satisfying \(\braces{R_d} \cup \braces{R_d \exp(\pi \hat{s}} \) for \( s \in \braces{e_1, e_2, e_3} \) .
The desired equilibrium is \( e_R = 0 \) and \( A = 0\).
\subsection{Proof of~\Cref{prop:error_dyn}}\label{proof:error_dyn}
From the kinematics~\cref{eqn:Rdot} and noting that \( \dot{R}_d = 0 \) the time derivative of \( R_d^T R \) is given as
\begin{gather*}
	\diff{}{t} \parenth{R_d^T R} = R_d^T R \hat{e}_\Omega \, .
\end{gather*}
Applying this to the time derivative of~\cref{eqn:A} gives
\begin{gather*}
	\diff{}{t} A = -\frac{1}{2} \tr{G R_d^T R \hat{e}_\Omega} \, .
\end{gather*}
Applying~\cref{eqn:hat1} into this shows~\cref{eqn:A_dot}.
Next, the time derivative of the repulsive error function is given by
\begin{gather*}
	\diff{}{t} B = \frac{r_i^T \parenth{\hat{\Omega} R^T} v_i}{\alpha \parenth{r_i^T R^T v_i - \cos \theta}} \, .
\end{gather*}
Using the scalar triple product, given by~\cref{eqn:STP}, one can reduce this to~\cref{eqn:B_dot}.
The time derivative of the attractive attitude error vector, \( e_{R_A} \), is given by
\begin{gather*}
	\diff{}{t} e_{R_A} = \frac{1}{2} \parenth{\hat{e}_\Omega R^T R_d G + (R^T R_d G)^T \hat{e}_\Omega}^\vee \, .
\end{gather*}
Using the hat map property given in~\cref{eqn:xAAx} this is further reduced to~\cref{eqn:eRA_dot,eqn:E}.
